### 2016-01-29 intro.Rmd
ENH:  Read-through.

Machine learning and pattern recognition techniques have seen increased application
for various medical image analysis workflows (see, for example, the annual
Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
Image Computing and Computer-Aided Intervention international meeting [@mlmi2015]).
Popular techniques such as support vector machines and neural networks have been applied
successfully to clinically relevant imaging tasks such as supervised image segmentation
(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
Facilitating the current employment of such techniques are the number of available imaging
data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
SciPy [@scipy] and the R project for statistical computing [@R] and their associated add-on
toolkits.

Random forests [@breiman2001] is a popular machine learning technique that has
demonstrated significant utitility for supervised segmentation tasks (e.g.,
normal human brain segmentation) and other computer vision applications
(e.g., human gait detection [@viola2005]).
In the context of neuropathology, random forest-based paradigms have been employed in the
delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
latter random forest approaches for brain tumor segmentation have performed well in recent
international competitions.  In response to the lack of objective comparisons between
segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
2012 [@Menze:2015aa] and has continued every year since under the auspices of the International Conference
of Medical Image Computing and Computer Assisted Interventions (MICCAI).

Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
that are built from training data.  Once constructed, data to be classified is "pushed"
through each decision tree resulting in a single classification "vote" per tree.  These
votes are then be used for regression or classification of the data.  Although decision
trees had been extensively studied previously, the success of employing collections of
such weak learners for boosting machine learning performance
(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
data (i.e., "bagging") resulting in the current random forest framework.

In this work, we develop a concatenated random forest framework with a tailored feature image set
for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
the entire framework is provided as open source through the well-known open-source
ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the investigation of this work is
the additional public availability of the TBI cohorts thus permitting full reproducibility
of the results reported and discussed.

>>>>

Machine learning and pattern recognition techniques have seen increased application
for various medical image analysis workflows (see, for example, the annual
Workshop on Machine Learning in Medical Imaging held in conjunction with the Medical
Image Computing and Computer-Aided Intervention (MICCAI) international meeting [@mlmi2015]).
Popular techniques such as support vector machines and neural networks have been applied
successfully to clinically relevant imaging tasks such as supervised image segmentation
(e.g., [@Bauer:2011aa]) and diagnostic prediction (e.g., [@Tong:2014aa;@Liu:2013aa]).
Facilitating the current employment of such techniques are the number of available imaging
data sets [@Van-Horn:2014aa] and the public availability of data science packages such as
SciPy [@scipy] and the R project for statistical computing [@R] and their associated
extensions.

The random forests framework [@breiman2001] is a popular machine learning technique that has
demonstrated significant utitility for supervised segmentation tasks (e.g.,
normal human brain segmentation [@yi2009]) and other computer vision applications
(e.g., human gait detection [@viola2005]).
In the context of neuropathology, random forest-based paradigms have been employed in the
delineation of multiple sclerosis lesions [@geremia2011], stroke lesions [@Pustina:2016aa],
and brain tumors [@geremia2012;@bauer2012;@zikic2012;@Tustison:2015aa].  Of note, these
latter random forest approaches for brain tumor segmentation have performed well in recent
international competitions.  In response to the lack of objective comparisons between
segmentation algorithms, the Multimodal Brain Tumor Segmentation (BRATS) challenge was initiated in
2012 [@Menze:2015aa] and has continued every year since under the auspices of the MICCAI
conference.

Random forests are conceptually simple [@breiman2001].  They consist of ensembles of decision trees
that are built from training data.  Once constructed, data to be classified is "pushed"
through each decision tree resulting in a single classification "vote" per tree.  These
votes are then used for regression or classification of the data.  Although decision
trees had been extensively studied previously, the success of employing collections of
such weak learners for boosting machine learning performance
(e.g., AdaBoost [@schapire1990;@freund1997]) influenced the similarly sytled conglomeration
of decision trees into "forests" with randomized node optimization [@ho1995;@amit1997].
Finally, Breiman [@breiman2001] improved accuracy by random sampling of the training
data (i.e., "bagging") resulting in the current random forest technique.

In this work, we develop a concatenated random forest framework with a tailored contextual
feature image set (both spatial and intensity-based)
for segmenting white matter hyperintensities in traumatic brain injury cohorts.  Additionally,
the entire framework is provided publicly through the well-known open-source
ANTs[^i1] and ANTsR[^i2] toolkits.  Further motivating the research of this work is
the public availability of the imaging data thus permitting full reproducibility
of the results reported and discussed.
